### 2018-02-15 main.tex
Initial commit (copying roughly from failed overleaf)

\documentclass[12pt, english]{wlpeerj} % Just a convenient template
\usepackage{palatino}
\usepackage{subfiles}
\usepackage[english]{babel} 
\usepackage[comma]{natbib}
\bibliographystyle{chicago2}
\usepackage{url}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{float}
\usepackage{verbatim}
\usepackage{fancyvrb}
\usepackage{tabularx}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage[colorlinks=true, citecolor = blue, linkcolor = black]{hyperref}
\usepackage{environ}
\usepackage{marginnote}
\usepackage{lineno}
\usepackage{setspace}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\reversemarginpar
\singlespacing

\newcommand{\jd}[1]{{\bf \color{blue} JD: #1}}
\newcommand{\mk}[1]{{\bf \color{red} MK: #1}}
\newcommand{\bmb}[1]{{\bf \color{green} BMB: #1}}

\title{Is statistical significance a thing?}

\author[1]{Jonathan Dushoff}
\author[1]{Morgan P. Kain}
\author[1,2]{Ben M. Bolker?}

\affil[1]{Department of Biology, McMaster University, 1280 Main Street West, Hamilton, Ontario L8S 4K1 Canada}
\affil[2]{Department of Mathematics and Statistics, McMaster University, 1280 Main Street West, Hamilton, Ontario 
L8S 4L8 Canada}

\corrauthor[1]{Jonathan Dushoff}{dushoff@mcmaster.ca}

\begin{document}

\input{doc.tex}

\newpage
\singlespacing

\bibliography{stat_signif}

\end{document}