### 2018-05-15 main.tex
Slight progress (for lots of effort)

\documentclass[12pt, english]{wlpeerj} % Just a convenient template
\usepackage{palatino}
\usepackage{subfiles}
\usepackage[english]{babel} 
\usepackage[comma]{natbib}
\bibliographystyle{unsrtnat}
\usepackage{url}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{float}
\usepackage{verbatim}
\usepackage{fancyvrb}
\usepackage{tabularx}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage[colorlinks=true, citecolor = purple, linkcolor = purple]{hyperref}
\usepackage{environ}
\usepackage{marginnote}
\usepackage{lineno}
\usepackage{setspace}
\usepackage{framed}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\reversemarginpar
\singlespacing

%% This whole effort was a disaster!
%% \usepackage[left=2.0in]{geometry} % For reasonable spacing with todonotes
% \usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
% \newcommand{\note}[2]{{\todo[linecolor=#1,backgroundcolor=#1!25,bordercolor=#1]{#2}{}}}

\newcommand{\note}[2]{{\bf \color{#1} #2}}
\newcommand{\jd}[1]{\note{blue}{JD: #1}}
\newcommand{\mk}[1]{\note{red}{MK: #1}}
\newcommand{\bmb}[1]{\note{green}{BMB: #1}}

\title{I can see clearly now: reinterpreting statistical significance
}

\author[1]{Jonathan Dushoff}
\author[1]{Morgan P. Kain}
\author[1,2]{Ben M. Bolker}

\affil[1]{Department of Biology, McMaster University, 1280 Main Street West, Hamilton, Ontario L8S 4K1 Canada}
\affil[2]{Department of Mathematics and Statistics, McMaster University, 1280 Main Street West, Hamilton, Ontario 
L8S 4L8 Canada}

\corrauthor[1]{Jonathan Dushoff}{dushoff@mcmaster.ca}

\begin{document}

\input{doc.tex}

\newpage
\singlespacing

\bibliography{stat_signif}

\end{document}