### 2019-01-07 table.tex
Cleaning

\newcommand{\ourcomment}[1]{[\emph{#1}]}
\begin{table}
\setlength\tabcolsep{1cm}.
\begin{tabular}{p{7.0cm}p{7.0cm}}
\textbf{Language from published articles} & \textbf{Rewritten using ``clarity''} \\
\hline\\

\multicolumn{2}{c}{\emph{\textbf{Accepting the null hypothesis ($\boldsymbol{p > 0.05 \nRightarrow}$ no effect)}}} \\
\hline

Toxins accumulate after acute exposure but have no effect on behaviour
& Toxins accumulate after acute exposure but their effects on behaviour are statistically unclear
\\

\rowcolor{vlg}
There was no effect of elevated carbon dioxide on reproductive behaviors 
& The effect of elevated carbon dioxide on reproductive behaviors was statistically unclear
\\

The finding that species richness showed no significant relationship with the area of available habitat is surprising because richness is usually strongly influenced by landscape context 
& Although species richness is usually strongly influenced by landscape context, we were unable to find a statistically clear relationship in this study
\\ \\

\multicolumn{2}{c}{\emph{\textbf{Inferring weak effects from large \pvals}} \citep{WassersteinandLazar2016}}
\\
\hline
%% left ``P values'' unchanged here (not \pvals) since it's a quotation
\ldots\ differences between treatment and control groups were nonsignificant, with P values of at least 0.3, and most in the range $0.7 \leq P \leq 0.9$.
& \ldots\ differences between treatment and control groups were not statistically clear (all $P > 0.05$) \ourcomment{since smallness is no longer implied the authors might now think of adding confidence intervals.}
\\ \\

\multicolumn{2}{c}{\emph{\textbf{The difference between ``clear'' and ``not clear'' is not clear}} \citep{GelmanandStern2006}}
\\
\hline

This correlation was significant in males ($\rho=0.35$, P \textless 0.05) but not females ($\rho=0.35$, NS). \ldots \ourcomment{The authors later write as though they have demonstrated a difference between males and females}
& Although males and females show the same correlation coefficient ($\rho=0.35$), the sign of the coefficient is statistically clear only in males  \ldots \ourcomment{Again, this phrasing may suggest to the authors that confidence intervals are called for.}
\\
\rowcolor{vlg}
\ldots risk of low BMD [bone mineral density] remained greater in HCV-coinfected women versus women with HIV alone
(adjusted OR 2.99, 95\% CI 1.33–6.74), but no association was found between HCV coinfection and low BMD in men 
(adjusted OR 1.26, 95\% CI 0.75–2.10). \ldots 
The precise mechanisms for the association between viral hepatitis and low BMD in HIV-infected women but not men remain unclear.
& 
\ldots  risk of low BMD [bone mineral density] remained greater in HCV-coinfected women versus women with HIV alone
(adjusted OR 2.99, 95\% CI 1.33–6.74), but the association between HCV coinfection and low BMD in men was not
statistically clear (adjusted OR 1.26, 95\% CI 0.75–2.10). \ldots  Pursuing biological differences between women and men in the effect of HIV on BMD would be premature given these results.
\\
\end{tabular}
\captionof{table}{Examples of misleading language in peer-reviewed papers (citations available by request), and revisions using our proposed  language of clarity.}
\label{quotetab}
\end{table}